Planarity and Street Network Representation in Urban Form Analysis

### 2018-02-08 main.tex
update figs

Models of street networks underlie research in urban travel behavior, accessibility, design patterns, and morphology. These models are commonly defined as planar, meaning that they can be represented in two dimensions without any underpasses or overpasses. However, real-world urban street networks exist in three-dimensional space and frequently feature grade separation such as bridges and tunnels: planar simplifications can be useful but might affect the results of real-world street network analysis. Further, a city's extent of nonplanarity could describe its infrastructure and urbanization. This study measures the nonplanarity of drivable and walkable street networks in the centers of 50 cities worldwide, then examines the variation of nonplanarity across a single city. It develops two new indicators: the Spatial Planarity Ratio and the Edge Length Ratio. While some street networks are approximately planar, we empirically quantify how planar simplifications can inconsistently but drastically misrepresent routing, connectivity, intersection counts/densities, and block lengths.

@@

\begin{figure}[htbp]
	\includegraphics[width=0.48\textwidth]{figures_tables/regression_phi_split.png}
	\caption{Log-log plot of $\lambda$ vs $\phi$ by network type with simple regression lines.}
	\label{fig:regression_split}
\end{figure}

@@

Figure \ref{fig:regression_split} depicts the relationship between the $\phi$ and $\lambda$ indicators across all 50 cities for both network types, as a log-log plot. The indicators' log-log relationship is linear, positive, and very strong (drivable $r^2=0.98$ and walkable $r^2=0.99$). The coefficients of determination suggest that these two indicators unsurprisingly provide redundant statistical information about the extent to which a network is spatially planar. However, each assesses different implications of this bias for measuring the urban form.

>>>>

Models of street networks underlie research in urban travel behavior, accessibility, design patterns, and morphology. These models are commonly defined as planar, meaning that they can be represented in two dimensions without any underpasses or overpasses. However, real-world urban street networks exist in three-dimensional space and frequently feature grade separation such as bridges and tunnels: planar simplifications can be useful but might affect the results of real-world street network analysis. Further, a city's extent of nonplanarity could describe its infrastructure and urbanization. This study measures the nonplanarity of drivable and walkable street networks in the centers of 50 cities worldwide, then examines the variation of nonplanarity across a single city. It develops two new indicators: the Spatial Planarity Ratio and the Edge Length Ratio. While some street networks are approximately planar, we empirically quantify how planar models can inconsistently but drastically misrepresent intersection counts/densities, block lengths, routing, and connectivity.

@@

\begin{figure}[htbp]
	\includegraphics[width=0.48\textwidth]{figures_tables/regression_lambda_phi_linear.png}
	\caption{Plot of $\lambda$ vs $\phi$ by network type with simple regression lines (dashed=drive, dotted=walk).}
	\label{fig:regression_lambda_phi}
\end{figure}

@@

Figure \ref{fig:regression_lambda_phi} depicts the relationship between $\phi$ and $\lambda$ across all 50 cities for both network types: it is linear, positive, and very strong (drivable $r^2=0.985$ and walkable $r^2=0.994$). The coefficients of determination suggest that these two indicators unsurprisingly provide redundant statistical information about the extent to which a network is spatially planar. However, each assesses different implications of this bias for measuring the urban form.
